---
title: "TP- PAM"
author: "Madalina Ciortan"
date: "December 17, 2016"
output: html_document
---

## TP: Substitution matrices

Read data from the file (note that columns names and row names should be discarded for subsequent computation and that the values given should be divided by 10000).


```{r}
setwd("D:\\workspace\\bioinformatics\\pam")
P1<-read.table('pam1.mat', sep = "\t", header=T, colClasses=c("character",rep("integer",20)))  
P1 <- P1[1:20,2:21]
P1 <-as.matrix(P1)

P1 <- P1/10000
#Print P1 diag
diag(P1)
```

Compute the probability matrix corresponding to any PAM distance n (corresponding to an evolutionary distance of n PAM, i.e. the PAMn probability matrix). 

```{r}
mypower <- function(x,n){
  p<-x
  prod<-diag(nrow(x))
  while ( n > 0 ){
    if(n %% 2 == 1) {
      prod <- prod %*% p
    } 
    n <- floor(n/2)
    p <- p %*% p
  }
  prod
}
P10 <- mypower(P1, 10)
diag(P10)

```

Compute the scoring matrix corresponding to your PAMn matrix.

score(i,j) = log(Pn / fj)

```{r}
freq <-read.table('aafreq.mat', sep = "\t", header=F)  
f <- freq$V2

score <- function(x, i, j) {
  log10(x/f[i])
}


scoringMatrix <- function(mat) {
  matrix(mapply(score, mat, row(mat), col(mat)), nrow = nrow(mat))
}

diag(scoringMatrix(P10))
```


Check the convergence of the values in the columns of the PAM probability matrix when n becomes large: 
These values should approach the amino acid frequencies. 

```{r}
convergence <- c (sum (f - diag(mypower(P1, 1500))), 
                  sum (f - diag(mypower(P1, 1650))), 
                  sum (f - diag(mypower(P1, 1800))), 
                  sum (f - diag(mypower(P1, 2000))),
                  sum (f - diag(mypower(P1, 2150))),
                  sum (f - diag(mypower(P1, 2300))))
                  
hist(convergence)
```


Plot the relation between the evolutionary distance (in PAM unit) and the percentage of identity between the sequences 

```{r}
nIndexes <- seq(1,1500, 20) # take n values at 20 Pam distance from 1 to 1500
distance <- function (value) {
  sum(diag(mypower(P1, value)))
}

hist(sapply(nIndexes, distance), nclass = 50)
```

